### 2017-08-22 01_actor_careers/05_report.Rmd
fixes and theme added

---
title: "Films: Age & Gender"
output: 
html_document:
code_folding: hide
---

# Questions?

Let's say we had a couple of questions about actors and directors:

> Do actors and directors improve with age?

> Are female actors generally cast  than their male counterparts?

To answer these questions we're going to need some data.

# Housekeeping

Let's set ourselves up with some packages, themes and utility functions.

```{r}
# Packages
library(readr)
library(dplyr)
library(tidyr)
library(gender)
library(stringr)
library(lubridate)
library(highcharter)

# Rmarkdown Options
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)

# Convert an R vector to string JS vector
convert_js_vec = function(vec) {
  js_string = vec %>% round(4) %>% paste(collapse = ",")
  return(JS( paste0("[", js_string, "]") ))
}

# Highcharts theme
source()
tmbish = high_tmbish()
```

# Data

I started with imdb user **hagennelson's** list of [top 1000 actors](http://www.imdb.com/list/ls058011111/) and other imdb user **earthanime's** list of [top 250 directors](http://www.imdb.com/list/ls062419846/). Both lists span across generations, from the ancient (Ethel Barrymore and Alfred Hitchcock) to the modern (Chloë Grace Moretz and Christopher Nolan).

I then developed code (which I'll cover in a different post) to scrape data from the rotten tomatoes pages of the above individuals' films. The final dataset includes **12,258** films and **1,233** indviduals.

Altough we'll only use a few of these variables I brought in a few extras for some future investigations:

Variable | Description
-------- | -----------
url | URL extension on rotten tomatoes website
year | Film production year
title | Film title
reviews | Count of formal critic reviews of film
tomatometer | Percentage of favourable reviews
av_rating | Average rating of film out of 10
audience | The non-critic plebian equivolent of the tomatometer
box_office | The box-office revenue
director | The film's director
actor_1:actor_6 | The 6 top actors according to rotten tomatoes internal ranking (seems like screen time)

## Film Data

We'll load and take a peek at the film dataset:

```{r}

films = read_csv("https://raw.githubusercontent.com/TMBish/lab_posts/master/01_actor_careers/00_full_film_data.csv") %>% filter(!is.na(title))

```

## Individual Data

Now we'll load the individual data.

Unfortunately, because the Rotten Tomatoes devs are a bunch of politically correct SJWs, gender isn't provided on their film bio pages so we'll have to be creative. The `gender` package in R allows you to infer a person's gender from their first name and DOB. The result is a probabilistic best guess of that individuals most likely gender given their first name and DOB.

```{r}

# library(gender)
# library(stringr)
# library(lubridate)
# 
# film_folk = read_csv("https://raw.githubusercontent.com/TMBish/lab_posts/master/01_actor_careers/00_actors_directors.csv")
# 
# # Clean some data quality issues on the DOB
# film_folk[film_folk$dob < as.Date("1880-01-01"), "dob"] = as.Date("1880-01-01")
# 
# # Grep out first name and dob year for the gender function
# film_folk =
#   film_folk %>%
#   mutate(
#     first_name = str_extract(name, "(?i)[a-z]+(?=\\s)"),
#     dob_year = year(dob)
#   )
# 
# # Distinct name, dob combos and predict gender
# genders =
#   film_folk %>%
#   distinct(first_name, dob_year) %>%
#   gender_df(name_col = "first_name", year_col = "dob_year") %>%
#   select("first_name" = name, gender, "dob_year" = year_min)

```

We should sense check the gender mapping. Let's take a random 10 people to check is working as desired:

```{r}
set.seed(123)

genders %>%
  sample_n(10) %>%
  knitr::kable(caption = "Sense Checking the Gender Mapping")

```

Things get dicey with a name like Sandy; however, the package obviously knows that most Sandys born in 1937 were female so we'll have to trust it here. Otherwise the list looks pretty, pretty, pretty good.

```{r}
# Join back the gender back to the main data
film_folk = 
  film_folk %>%
  left_join(genders)

```

# Q1. Gender Gap

To answer this question we'll first need to calculate a person's age at the time of their appearance in a film. We'll then

```{r}

# Chain a bunch of data wrangling operations
volume_data =
  films %>% 
  select(title, year, tomatometer, director, actor_1:actor_6) %>%
  gather("role", "name", director, actor_1:actor_6) %>%
  mutate(role = ifelse(role == "director", "Director", "Actor")) %>%
  inner_join(film_folk) %>%
  mutate(
    age_at_production = year - dob_year
  ) %>%
  filter(
    between(age_at_production, 10,100),
    !is.na(gender)
  ) %>%
  select(name, gender, dob_year, age_at_production, year, role, tomatometer) %>%
  mutate(
    category = case_when(
      role == "Director" ~ "director",
      gender == "male" ~ "male_actor",
      TRUE ~ "female_actor"
    )
  ) %>% 
  group_by(category, age_at_production) %>%
  summarise(
    films = n()
  ) %>%
  group_by(category) %>%
  mutate(
    percentage = films / sum(films)
  )

```

<div id = "age_series_chart">
```{r}

# Shape data for presentation
chart_data = 
  volume_data %>%
  select(-films) %>%
  spread(category, percentage) %>%
  mutate(
    director = ifelse(!is.na(director), director, 0),
    female_actor = ifelse(!is.na(female_actor), female_actor, 0),
    male_actor = ifelse(!is.na(male_actor), male_actor, 0),
    director_cdf = cumsum(director),
    female_actor_cdf = cumsum(female_actor),
    male_actor_cdf = cumsum(male_actor)
  )

# For the javascript client side interactivity on the age series chart
female_cdf = chart_data$female_actor_cdf %>% convert_js_vec()
female_pdf = chart_data$female_actor %>% convert_js_vec()
male_cdf = chart_data$male_actor_cdf %>% convert_js_vec()
male_pdf = chart_data$male_actor %>% convert_js_vec()
dir_cdf = chart_data$director_cdf %>% convert_js_vec()
dir_pdf = chart_data$director %>% convert_js_vec()

# Create chart
age_dist_chart = 
  highchart() %>%
  hc_add_theme(tmbish) %>%
  hc_chart(
    animation = list(duration = 2000)
  ) %>%
  hc_xAxis(
    categories = chart_data$age_at_production,
    title = list(text = "Age at Production")
  ) %>%
  hc_yAxis(
    title = list(text = "Probablity Density"),
    labels = list(formatter = JS("function(){return(Math.round(this.value * 1000) / 10 + '%')}"))
  ) %>%
  hc_title(text = "Female actors") %>%
  hc_subtitle(text = "Comparing male actors, (male) directors and female actors") %>%
  hc_add_series(
    name = "Director", 
    type = "spline",
    data = chart_data$director,
    marker = list(enabled = FALSE)
  ) %>%
  hc_add_series(
    name = "Male Actor",
    type = "spline",
    data = chart_data$male_actor,
    marker = list(enabled = FALSE)
  ) %>%
  hc_add_series(
    name = "Female Actor",
    type = "area",
    data = chart_data$female_actor,
    zIndex = -10,
    marker = list(enabled = FALSE),
    fillOpacity = 0.5
  )

age_dist_chart
```
</div>

<button class="button secondary outline" id = "update_age_series"> Show Cumulative Distribution </button>

# Q2. Acting and Directing as a skill

> Hypothesis 2: acting and directing skills aren't tied to a person's physical attributes so shouldn't experience deterioration in age. In fact you would expect these artists to improve their skills with age.

## Measuring Performance

To investigate our hypothesis we must make a couple of assumptions. Since we don't have an individual actor / director performance (like we have in sports) we'll have to assume an actor or director can have a tangible affect on the quality of the movie they appear in; there are obviously many other factors but this seems like a relatively safe assumption.

Further - for actors - we'll have to limit the dataset to actors that played an important role in the film. To implement an `actor performance == film quality` approach the actors need to be prominent enough so as to be able to affect the overall quality of the film. So while Michael Cera's cameo in **This is the End** was high quality it can't be said that he influenced the rotten tomatoes rating of the film on his own.

<iframe width="560" height="315" src="https://www.youtube.com/embed/qIKPJlKHKxg" frameborder="0" allowfullscreen></iframe>

After looking through a few rotten tomatoes pages it seems limiting to the top 3 actors in a given film is a reasonable cut-off for influncial performances.

> INSERT EXAMPLES FROM A FEW ROTTEN TOMATOES PAGES

## Methodology

For every person in the dataset we follow the following recipe:

1. Produce a dataset for each actor with films that:
+ We have a tomatometer score for
+ And the actor is in the top 3 of rotten tomatoes credits
2. Calculate the `mean` tomatometer for all of this person's films;
3. Calculate the age the person was at the release of each film;
4. For every film calculate what `percentage %` this tomato rating is above or below this person's `mean` (called a tomato index)

Producing a dataset like this:

```{r echo=FALSE}
films %>% 
  select(title, year, tomatometer, director, actor_1:actor_3) %>%
  gather("role", "name", director, actor_1:actor_3) %>%
  inner_join(film_folk) %>%
  filter(
    name == "Denzel Washington"
  ) %>%
  mutate(
    age_at_production = year - dob_year
  ) %>%
  filter(between(age_at_production, 15,85)) %>%
  select(title, name, year, gender, age_at_production, tomatometer) %>%
  group_by(name) %>% 
  mutate(
    average_tomato  = round(mean(tomatometer))
  ) %>% 
  ungroup() %>%
  mutate(
    tomato_index = tomatometer / average_tomato
  ) %>%
  arrange(age_at_production) %>%
  select(title, name, year, age_at_production, tomatometer, average_tomato, tomato_index) %>%
  mutate(index = scales::percent(tomato_index)) %>%
  sample_n(10) %>%
  knitr::kable()

```

## Results

```{r}
# 
rating_set =
  films %>% 
  select(title, year, tomatometer, director, actor_1:actor_3) %>%
  gather("role", "name", director, actor_1:actor_3) %>%
  inner_join(film_folk) %>%
  filter(!is.na(gender)) %>%
  mutate(
    age_at_production = year - dob_year,
    category = case_when(
      role == "director" ~ "director",
      gender == "male" ~ "male_actor",
      TRUE ~ "female_actor"
    )
  ) %>%
  filter(between(age_at_production, 15,85)) %>%
  select(name, gender, age_at_production, category, tomatometer) %>%
  group_by(name) %>% 
  mutate(
    average_tomato  = median(tomatometer),
    films = n()
  ) %>% ungroup() %>%
  filter(films >= 5)

rating_volumes = 
  rating_set %>%
  group_by(age_at_production) %>%
  summarise(films = -1*n())

rating_scores = 
  rating_set %>%
  mutate(
    tomato_index = tomatometer / average_tomato) %>%
  group_by(category, age_at_production) %>%
  summarise(rating = median(tomato_index)) %>%
  spread(category, rating)

# LOESS Trend Line
loess_fit = loess(value ~ age_at_production , 
                  data = rating_scores %>% gather("role", "value", -age_at_production)
                  # %>% group_by(age_at_production) %>% summarise(value = mean(value, na.rm = TRUE))
)

predictions = loess_fit %>% predict(seq(15,85))

volume_chart = 
  highchart() %>%
  hc_add_theme(tmbish) %>%
  hc_chart(height = "300") %>%
  hc_xAxis(categories = rating_volumes$age_at_production, opposite = TRUE) %>%
  hc_yAxis(
    title = list(text = "Total Films in Dataset"),
    labels = list(formatter = JS("function(){return(-1 * this.value)}"))    
    ) %>%
  hc_title(text = "") %>%
  hc_add_series(
    name = "Volumes",
    type = "area",
    data = rating_volumes$films,
    marker = list(enabled = FALSE)
  ) %>% 
  hc_legend(enabled = FALSE)

rating_chart = 
  highchart() %>%
  hc_add_theme(tmbish) %>%
  hc_chart(
    animation = list(duration = 2000)
  ) %>%
  hc_xAxis(
    categories = rating_scores$age_at_production,
    title = list(text = "Age at Production")
  ) %>%
  hc_yAxis(
    title = list(text = "Probablity Density"),
    min = 0.6
  ) %>%
  hc_title(text = "Female actors") %>%
  hc_subtitle(text = "Comparing male actors, (male) directors and female actors") %>%
  hc_add_series(
    name = "Director",
    type = "scatter",
    data = rating_scores$director,
    marker = list(radius = 3, symbol = "circle")
  ) %>%
  hc_add_series(
    name = "Male Actor",
    type = "scatter",
    data = rating_scores$male_actor,
    marker = list(radius = 3, symbol = "circle")
  ) %>%
  hc_add_series(
    name = "Female Actor",
    type = "scatter",
    data = rating_scores$female_actor,
    marker = list(radius = 3, symbol = "circle")
  ) %>%
  hc_add_series(
    name = "Trend",
    type = "spline",
    data = predictions,
    color = "#000000",
    marker = list(enabled = FALSE)
  )

```

```{r echo = FALSE}
chart
```

```{r echo = FALSE}
volume_chart
```

<script>

$(document).ready(function() {

var show_cdf = false;

$("#update_age_series").click(function(){

var chart=$("#age_series_chart").find(".highchart").highcharts();

if (show_cdf) {

chart.yAxis[0].setExtremes(0,null);
chart.yAxis[0].setTitle({text:"Probability Density", redraw:false});
dir_pdf = `r dir_pdf`;
female_pdf = `r female_pdf`;
male_pdf = `r male_pdf`;
chart.series[0].setData(dir_pdf);
chart.series[1].setData(male_pdf);
chart.series[2].setData(female_pdf);
$("#update_age_series").html("Show Cumulative Distribution");

} else {

chart.yAxis[0].setExtremes(0,1);
chart.yAxis[0].setTitle({text:"Cumulative Probability", redraw:false});
dir_cdf = `r dir_cdf`;
female_cdf = `r female_cdf`;
male_cdf = `r male_cdf`;
chart.series[0].setData(dir_cdf);
chart.series[1].setData(male_cdf);
chart.series[2].setData(female_cdf);

$("#update_age_series").html("Show Density");

};

show_cdf = !show_cdf

}); 
});

</script>