### 2017-08-30 01_actor_careers/05_report.Rmd
updated theme

---
title: "The prime of their life - the halflife of acting talent"
output: 
html_document:
code_folding: hide
---

> insert an image

Much analysis has been completed regarding the performing peak of athletes and sportspeople. Statistically, swimmers peak at around 24, marathon runners between 30-31, gymnasts at a mere 18, NBA players generally peak between 26-29, professional hockey players between 27-28, and chess players at 30-35.

Does the idea of a performing prime extend to artists, specifically actors and directors? There are obvious reasons to be skeptical. Importantly there's no major physical component to acting that would decay as the body does, and unlike chess you wouldn't think there's the need for an exreme state of concentration and cognition which might be unreachable at older ages.

There are significant confounding factors that might prevent us from answering this question difinitively (which I'll discuss later) but let's dig into the data.

```{r echo = FALSE, warning=FALSE, message=FALSE}
# Let's set ourselves up with some packages, themes and utility functions.

# Packages
library(readr)
library(dplyr)
library(tidyr)
library(gender)
library(stringr)
library(lubridate)
library(highcharter)
library(purrr)
library(htmltools)

# Rmarkdown Options
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)

# For reproducibility
set.seed(123)

# Convert an R vector to string JS vector
convert_js_vec = function(vec) {
  js_string = vec %>% round(4) %>% paste(collapse = ",")
  return(JS( paste0("[", js_string, "]") ))
}

# Highcharts theme
source("https://raw.githubusercontent.com/TMBish/lab_posts/master/highcharts_theme.R")
tmbish = high_tmbish()
```

# Data

To curate a data set for anlaysis I started with imdb user **hagennelson's** list of [top 1000 actors](http://www.imdb.com/list/ls058011111/) and imdb user **earthanime's** list of [top 250 directors](http://www.imdb.com/list/ls062419846/). Both lists span across generations, from the ancient (Ethel Barrymore and Alfred Hitchcock) to the modern (Chloë Grace Moretz and Christopher Nolan).

I then developed code (available on GitHub) to scrape data from the rotten tomatoes pages of the above individuals' films. The final dataset includes **12,258** films and **1,233** indviduals. Although we'll only use a few of these variables, I brought in a few extras for some future investigations:

Variable | Description
------------ | ---------------------
url | URL extension on rotten tomatoes website
year | Film production year
title | Film title
reviews | Count of formal critic reviews of film
tomatometer | Percentage of favourable reviews
av_rating | Average rating of film out of 10
audience | The non-critic plebian equivolent of the tomatometer
box_office | The box-office revenue
director | The film's director
actor_1:actor_6 | The 6 top actors according to rotten tomatoes internal ranking (seemingly ordered by screen time)

## Film Data

Let's have a look at the films data set
```{r}
# Load in films data from Github Repo
films = read_csv("https://raw.githubusercontent.com/TMBish/lab_posts/master/01_actor_careers/00_full_film_data.csv") %>% filter(!is.na(title))

films %>% 
  select(-url) %>%
  sample_n(5) %>%
  knitr::kable()

```

## Individual Data

Now we'll load the individual data.

Unfortunately, because the Rotten Tomatoes development team are a bunch of politically correct SJWs, gender isn't provided on their film bio pages so we'll have to be creative. The `gender` R package allows you to infer a person's gender from their first name and DOB. The result is a probabilistic estimate of that individuals most likely gender given their first name and DOB.

```{r}

library(gender)
library(stringr)
library(lubridate)

film_folk = read_csv("https://raw.githubusercontent.com/TMBish/lab_posts/master/01_actor_careers/00_actors_directors.csv")

# Clean some data quality issues on the DOB
film_folk[film_folk$dob < as.Date("1880-01-01"), "dob"] = as.Date("1880-01-01")

# Grep out first name and dob year for the gender function
film_folk = film_folk %>%
              mutate(
                first_name = str_extract(name, "(?i)[a-z]+(?=\\s)"),
                dob_year = year(dob)
              )

# Distinct name, dob combos and predict gender
genders = film_folk %>%
            distinct(first_name, dob_year) %>%
            gender_df(name_col = "first_name", year_col = "dob_year") %>%
            select("first_name" = name, gender, "dob_year" = year_min)

```

We should sense check the gender mapping to make sure the package is making logical inferences. Let's take a random 10 person sample:

```{r}
genders %>%
  sample_n(10) %>%
  knitr::kable()
```

So generally this is looking pretty, pretty good. Things get dicey with unisex names like Sandy; however, the package must know that most Sandys born in 1937 were female so we'll have to trust it here. Finally we'll need to join this information back to the individual data set.

```{r}
# Join back the gender back to the main data
film_folk = 
  film_folk %>%
  left_join(genders)

```

# What's the deal with female directors?

To add a layer to this analysis I wanted group people by gender and film production role (actor vs director). So we think we'd be working with 4 groups, however, I was suprised to see that of our 250 top directors only 25 were female. A consequence of history but appaling nonetheless.

This got me thinking about the rest of the dataset. Given our **12,258** film dataset isn't limited to director credits in our top 250 (films in the set are contributed by both actor and director credits) we can get a bit of a better idea of the broader gender split for film directors. I've captured a lot of the key numbers in the following chart.

```{r}
# Establish a persistent dataset of interest
master = films %>%
        select(title, year, tomatometer, director:actor_6) %>%
        gather("role", "name", director:actor_6) %>%
        inner_join(film_folk) %>%
        select(-first_name) %>%
        mutate(
          role_type = ifelse(role == "director", "Director", "Actor"),
          age_at_production = year - dob_year,
          category = case_when(
            role_type == "Director" ~ "director",
            gender == "male" ~ "male_actor",
            gender == "female" ~ "female_actor",
            TRUE ~ "other"
          )
        )

# Directors

# Data for known individuals
known_summary = 
  master %>%
  group_by(role_type, gender) %>%
  summarise(
    films = n(),
    distinct_people = n_distinct(name)
  )

# Unknown directors
unknown_set = films %>%
          select("name" = director) %>%
          anti_join(film_folk) %>%
          mutate(
            first_name = str_extract(name, "(?i)[a-z]+(?=\\s)"),
            min_year = 1900, max_year = 1980
          ) %>%
          filter(!is.na(first_name))

unknown_genders = 
  unknown_set %>% 
  gender_df(name_col = "first_name", year_col = c("min_year", "max_year")) %>%
  select("first_name" = name, gender)

unknown_summary = 
  unknown_set %>%
  inner_join(unknown_genders) %>%
  mutate(role_type = "Director (Non Top250)") %>%
  group_by(role_type, gender) %>%
  summarise(
    films = n(),
    distinct_people = n_distinct(name)
  )

chart_data = 
  known_summary %>%
  union_all(unknown_summary) %>%
  filter(!is.na(gender))

gender_props = list()

for (metric in c("films", "distinct_people")) {

  data = 
        chart_data %>%
        select_("role_type", "gender", metric) %>%
        spread_("gender", metric, fill = 0)

    base_chart = 
      highchart() %>%
      hc_chart(type = "column") %>%
      hc_add_theme(tmbish) %>%
      hc_xAxis(categories = data$role_type) %>%
      hc_plotOptions(
        column = list(
          stacking = 'percent',
          dataLabels = list(
            enabled = TRUE,
            color = "#FFF",
            style = list(textOutline = NULL, fontSize = "14px"),
            formatter = JS("function(){ return(Math.round(this.percentage) + '%')}")
          )
        )
      ) %>%
      hc_title(text = ifelse(metric=="films", "By Volume of Films", "By Distinct Individuals")) %>%
      hc_legend(align = "center", verticalAlign = "bottom", layout = "horizontal")

    for (i in c("female", "male")) {
      base_chart = base_chart %>% hc_add_series(name = i, data = data[,i] %>% unlist() %>% as.integer())
    }

    if (metric != "films") { base_chart  = base_chart %>% hc_legend(enabled=FALSE)}

  gender_props[[metric]] = base_chart

}

gender_props %>%
  hw_grid(ncol = 1, rowheight = "400px") %>%
  browsable()

```

> Insert finding graphic thing

So henceforth we'll treat directors as a single bloc and you can assume that they're male statistically speaking.

# I starred in or directed a film; how old am I likely to be?

Well, as it turns out, it depends. We'll group individuals in groups: directors (predominantly male as seen above), male actors, and female actors. For all these individuals we calculate their age at the time a film they were in was realeased and plot the densities within each group.

Switch between the denisty (proportion appearing at one speicific age) and cumulative distribution (proportion appearing at **or before** at specific age) views to better understand the data.

```{r}

# Chain a bunch of data wrangling operations
volume_data =
  master %>%
  filter(
    between(age_at_production, 10,100),
    !is.na(gender),
    category != "other"
  ) %>%
  group_by(category, age_at_production) %>%
  summarise(
    films = n()
  ) %>%
  group_by(category) %>%
  mutate(
    percentage = films / sum(films)
  )

```

<div id = "age_series_chart">
```{r}

# Shape data for presentation
chart_data = 
  volume_data %>%
  select(-films) %>%
  spread(category, percentage) %>%
  mutate(
    director = ifelse(!is.na(director), director, 0),
    female_actor = ifelse(!is.na(female_actor), female_actor, 0),
    male_actor = ifelse(!is.na(male_actor), male_actor, 0),
    director_cdf = cumsum(director),
    female_actor_cdf = cumsum(female_actor),
    male_actor_cdf = cumsum(male_actor)
  )

# For the javascript client side interactivity on the age series chart
female_cdf = chart_data$female_actor_cdf %>% convert_js_vec()
female_pdf = chart_data$female_actor %>% convert_js_vec()
male_cdf = chart_data$male_actor_cdf %>% convert_js_vec()
male_pdf = chart_data$male_actor %>% convert_js_vec()
dir_cdf = chart_data$director_cdf %>% convert_js_vec()
dir_pdf = chart_data$director %>% convert_js_vec()

# Create chart
age_dist_chart = 
  highchart() %>%
  hc_add_theme(tmbish) %>%
  hc_chart(
    animation = list(duration = 2000)
  ) %>%
  hc_xAxis(
    categories = chart_data$age_at_production,
    title = list(text = "Age at Production")
  ) %>%
  hc_yAxis(
    title = list(text = "Probablity Density"),
    labels = list(formatter = JS("function(){return(Math.round(this.value * 1000) / 10 + '%')}"))
  ) %>%
  hc_title(text = "Age at Production") %>%
  hc_subtitle(text = "Comparing the distribution of film professionals for age at the time of film productions") %>%
  hc_add_series(
    name = "Director", 
    type = "spline",
    data = chart_data$director,
    marker = list(enabled = FALSE)
  ) %>%
  hc_add_series(
    name = "Male Actor",
    type = "spline",
    data = chart_data$male_actor,
    marker = list(enabled = FALSE)
  ) %>%
  hc_add_series(
    name = "Female Actor",
    type = "spline",
    data = chart_data$female_actor,
    zIndex = -10
    #marker = list(enabled = FALSE)
    #fillOpacity = 0.5
  )

age_dist_chart
```
</div>

<button class="button secondary outline" id = "update_age_series"> Show Cumulative Distribution </button>

> Insert infographic thing

# The half life of acting talent

> Hypothesis 2: acting and directing skills aren't tied to a person's physical attributes so shouldn't experience deterioration in age. In fact you would expect these artists to improve their skills with age.

## Measuring Performance

To investigate this hypothesis we must make a couple of assumptions. Since we don't have individual actor / director performance metrics (like we have in sports) we'll have to assume an actor or director can have a material affect on the quality of the movie they appear in; there are obviously many other factors but this seems like a relatively safe assumption.

Further - for actors - we'll have to limit the dataset to actors that played an important role in the film. To implement an `actor performance == film quality` approach the actors need to be noteworthy enough to be able to affect the overall quality of the film. So while Michael Cera's cameo in **This is the End** was high quality it can't be said that he influenced the rotten tomatoes rating of the film on his own.

<iframe width="560" height="315" src="https://www.youtube.com/embed/qIKPJlKHKxg" frameborder="0" allowfullscreen></iframe>

After looking through a few rotten tomatoes pages it seems limiting to the top 3 actors in a given film is a reasonable cut-off for influncial performances.

> INSERT EXAMPLES FROM A FEW ROTTEN TOMATOES PAGES

## Methodology

There's a few other things we want to control for here. I don't want to plot raw film critical rating because we might be prone to idiosyncracies of groups or ages - high performing directors might generally wait until they feel experienced enough to direct their first film as an example.

Instead we'll plot, for each individual and film, how well they have perfomed relative to that individual's mean lifetime score (the average tomatometer of all that individual's films). This gives us a common scale on which to compare.

To illustrate this approach let's look at Denzel Washington's films. Denzel has a healthy mean tomatometer of 68%.

```{r echo=FALSE}
master %>% 
  filter(
    !(role %in% c("actor_4","actor_5", "actor_6")),
    name == "Denzel Washington",
    between(age_at_production, 15,85)
  ) %>%
  group_by(name) %>% 
  mutate(
    average_tomato  = round(mean(tomatometer))
  ) %>% 
  ungroup() %>%
  mutate(
    tomato_index = (tomatometer - average_tomato) / average_tomato
  ) %>%
  sample_n(10) %>%
  arrange(age_at_production) %>%
  select(title, name, year, age_at_production, tomatometer, average_tomato, tomato_index) %>%
  mutate(tomato_index = scales::percent(tomato_index)) %>%
  knitr::kable()

```

So based on this 10 film sample Denzel had a solid 40s where his films (of the likes **Hurricane** and **Courage Under Fire**) average a rotten tomatoes rating about 20% above his career average. He looks like he dropped off after 50 with a few flops like **Déjà Vu** (which I didn't mind) and **2 Guns**. Look I'm not calling him washed but...

## Results

So we use this approach and plot out the median tomato index by age at production, giving us an idea of the *average performance above or below a career average* at each age.

```{r}

# 
rating_set =
  master %>%
  filter(
    !(role %in% c("actor_4","actor_5", "actor_6")),
    between(age_at_production, 15,85)
  ) %>%
  group_by(name) %>% 
  mutate(
    average_tomato  = round(mean(tomatometer)),
    films = n()
  ) %>% 
  ungroup() %>%
  mutate(tomato_index = (tomatometer - average_tomato) / average_tomato) %>%
  filter(films >= 5)

rating_volumes = 
  rating_set %>%
  group_by(age_at_production) %>%
  summarise(films = -1*n())

rating_scores = 
  rating_set %>%
  group_by(category, age_at_production) %>%
  summarise(rating = median(tomato_index)) %>%
  spread(category, rating)

# LOESS Trend Line
loess_fit = loess(value ~ age_at_production , 
                  data = rating_scores %>% gather("role", "value", -age_at_production)
                  # %>% group_by(age_at_production) %>% summarise(value = mean(value, na.rm = TRUE))
)

predictions = loess_fit %>% predict(seq(15,85))

volume_chart = 
  highchart() %>%
  hc_add_theme(tmbish) %>%
  hc_chart(height = "300") %>%
  hc_xAxis(categories = rating_volumes$age_at_production, opposite = TRUE) %>%
  hc_yAxis(
    title = list(text = "Total Films in Dataset"),
    labels = list(formatter = JS("function(){return(-1 * this.value)}"))    
    ) %>%
  hc_title(text = "") %>%
  hc_add_series(
    name = "Volumes",
    type = "area",
    data = rating_volumes$films,
    marker = list(enabled = FALSE)
  ) %>% 
  hc_legend(enabled = FALSE)

rating_chart = 
  highchart() %>%
  hc_add_theme(tmbish) %>%
  hc_chart(
    animation = list(duration = 2000)
  ) %>%
  hc_xAxis(
    categories = rating_scores$age_at_production,
    title = list(text = "Age at Production")
  ) %>%
  hc_yAxis(
    title = list(text = "Probablity Density")
  ) %>%
  hc_title(text = "Female actors") %>%
  hc_subtitle(text = "Comparing male actors, (male) directors and female actors") %>%
  hc_add_series(
    name = "Director",
    type = "scatter",
    data = rating_scores$director,
    marker = list(radius = 3, symbol = "circle")
  ) %>%
  hc_add_series(
    name = "Male Actor",
    type = "scatter",
    data = rating_scores$male_actor,
    marker = list(radius = 3, symbol = "circle")
  ) %>%
  hc_add_series(
    name = "Female Actor",
    type = "scatter",
    data = rating_scores$female_actor,
    marker = list(radius = 3, symbol = "circle")
  ) %>%
  hc_add_series(
    name = "Trend",
    type = "spline",
    data = predictions,
    color = "#000000",
    marker = list(enabled = FALSE)
  )

```

```{r echo = FALSE}
rating_chart
```

```{r echo = FALSE}
volume_chart
```

# The prime suspects

```{r}

actors = c("Robert De Niro", "Harvey Keitel", "Christopher Walken", "Bruce Willis", "Susan Sarandon", "Jeff Bridges", "Robin Williams", "Julianne Moore", "Steve Buscemi", "John Malkovich", "Meryl Streep", "Robert Redford")

graphs = map(actors, function(actor) {

  bucket = 5

  actordata = 
    master %>%
    filter(
      name == actor,
      role %in% c("actor_1", "actor_2", "actor_3")
    ) %>%
    mutate(
      average_tomato  = round(mean(tomatometer)),
      tomato_index = (tomatometer - average_tomato) / average_tomato
    ) %>%
    mutate(age_bucket = round(age_at_production / bucket) * bucket) %>%
    group_by(age_bucket) %>%
    summarise(
      index = mean(tomato_index),
      films = n()) %>%
    filter(films >= 3)

  highchart() %>%
    hc_add_theme(tmbish) %>%
    hc_chart(
      animation = list(duration = 2000)
    ) %>%
    hc_xAxis(
      categories = actordata$age_bucket,
      title = list(text = "Age at Production")
    ) %>%
    hc_title(text = actor) %>%
    hc_add_series(
      name = "Performance Index",
      type = "spline",
      data = actordata$index
    ) %>%
    hc_legend(enabled = FALSE)

}) %>% hw_grid(ncol = 3, rowheight = 400) %>% browsable()

graphs

```

<script>

$(document).ready(function() {

var show_cdf = false;

$("#update_age_series").click(function(){

var chart=$("#age_series_chart").find(".highchart").highcharts();

if (show_cdf) {

chart.yAxis[0].setExtremes(0,null);
chart.yAxis[0].setTitle({text:"Probability Density", redraw:false});
dir_pdf = `r dir_pdf`;
female_pdf = `r female_pdf`;
male_pdf = `r male_pdf`;
chart.series[0].setData(dir_pdf);
chart.series[1].setData(male_pdf);
chart.series[2].setData(female_pdf);
$("#update_age_series").html("Show Cumulative Distribution");

} else {

chart.yAxis[0].setExtremes(0,1);
chart.yAxis[0].setTitle({text:"Cumulative Probability", redraw:false});
dir_cdf = `r dir_cdf`;
female_cdf = `r female_cdf`;
male_cdf = `r male_cdf`;
chart.series[0].setData(dir_cdf);
chart.series[1].setData(male_cdf);
chart.series[2].setData(female_cdf);

$("#update_age_series").html("Show Density");

};

show_cdf = !show_cdf

}); 
});

</script>