### 2017-09-17 02_rmarkdown-interactivity/post.Rmd
Progress on rmarkdown interactivity

---
title: "Adding Interactivity to Static Rmarkdown Documents"
author: ~
date: "`r Sys.time()`"
day: "`r format(Sys.Date(), '%d')`"
month: "`r format(Sys.Date(), '%b')`"
year: "`r format(Sys.Date(), '%Y')`"
slug: adding-rmarkdown-interactivity
categories: []
tags: ["R", "R Markdown", "Javascript", "Highcharts"]
highlight: "kate"
description: "Interactivity can be added to static HTML R Markdown outputs by using a couple of cool tricks."
img: "bulletin-edition-1"
---

# Overview

Interactivity in data analytics outputs is great for a number of reasons.

As an R User you have a number of different formats to write. It can become confusing about how each of these formats differ from one another, specifically when it comes to interactivity.

- **Shiny Apps** is the most interactive format with few limitations but requires a server to run.
- **HTML R markdown docs, slides and Notebooks** are nice comprimise.
- **Non-HTML Rmarkdown outputs** are the least interactive with text and plots (images) assembled in a static arrangement.

This post specifically focuses on option 2 and how it can be enhanced to be more like '

# Generate Data

```{r}
library(dplyr)
library(tidyr)

data <- data.frame(
  normal = rnorm(10000, 5, 2), # 1000 random from normal dist
  exponential = rexp(10000, 0.5), # 1000 random from exponential dist
  uniform = runif(10000, 0, 10) # 1000 random from uniform dist
)

# Data Manipulation
data %>%
  gather(distribution, value, normal:uniform) %>%
  mutate(value = value %>% round(1)) %>%
  filter(value >= 0, value <= 12) %>%
  group_by(distribution, value) %>%
  summarise(points = n()) %>%
  ungroup() %>% group_by(distribution) %>%
  mutate(density = points / sum(points)) %>%
  ungroup() %>%
  select(-points) -> density

density %>% head()

```

# Plot Data

```{r}

library(highcharter)

plot_data = density %>% spread(distribution, density)

  highchart() %>%
  hc_add_theme(hc_theme_monokai()) %>%
  hc_chart(
    animation = list(duration = 2000),
    height = 500
  ) %>%
  hc_xAxis(
    title = list(text = "Value")
  ) %>%
  hc_yAxis(
    title = list(text = "Probablity Density"),
    labels = list(formatter = JS("function(){return(Math.round(this.value * 1000) / 10 + '%')}"))
  ) %>%
  hc_title(text = "3 Probability Densities") %>%
  hc_add_series(
    name = "Normal", 
    type = "spline",
    data = plot_data %>% select(x = value, y = normal) %>% list_parse()
    #marker = list(enabled = FALSE)
  ) %>%
  hc_add_series(
    name = "Exponential",
    type = "spline",
    data = plot_data %>% select(x = value, y = exponential) %>% list_parse()
   # marker = list(enabled = FALSE)
  ) %>%
  hc_add_series(
    name = "Uniform",
    type = "spline",
    data = plot_data %>% select(x = value, y = uniform) %>% list_parse()
   # marker = list(enabled = FALSE)
  ) -> chart

chart

```

# Add Interactivity

```{r}

# Get CDF from density dataframe
get_dist = function(df, var, type) {

  if (type=="pdf") {

   df %>%
    filter(distribution == !!var) %>% # Tidy Evaluation
    mutate(
      js_pair = paste0("[", value, "," , density %>% round(6), "]")
    ) %>%
    pull(js_pair) %>% # Just found this function, handy!
    return()

  } else {

   df %>%
    filter(distribution == !!var) %>% # Tidy Evaluation
    arrange(value) %>%
    mutate(
      cdf = cumsum(density),
      js_pair = paste0("[", value, "," , cdf %>% round(4), "]")
    ) %>%
    pull(js_pair) %>% # Just found this function, handy!
    return()

  }

}

# Conver R char vector of x, y pairs to JS array string
js_stringify = function(vec) {

  vec %>%
    paste(collapse = ",") %>%
    paste0("[", . , "]") %>%
    JS() %>%
    return()

}

# Javascriptise PDF Data
nor_pdf = get_dist(density, "normal", "pdf") %>% js_stringify()
exp_pdf = get_dist(density, "exponential", "pdf") %>% js_stringify()
uni_pdf = get_dist(density, "uniform", "pdf") %>% js_stringify()

# Javascriptise CDF Data
nor_cdf = get_dist(density, "normal", "cdf") %>% js_stringify()
exp_cdf = get_dist(density, "exponential", "cdf") %>% js_stringify()
uni_cdf = get_dist(density, "uniform", "cdf") %>% js_stringify()

```

# In action

```{r}

  highchart(elementId = "densities_v2") %>%
  hc_add_theme(hc_theme_monokai()) %>%
  hc_chart(
    animation = list(duration = 2000),
    height = 500
  ) %>%
  hc_xAxis(
    #categories = plot_data$value,
    title = list(text = "Value")
  ) %>%
  hc_yAxis(
    title = list(text = "Probablity Density"),
    labels = list(formatter = JS("function(){return(Math.round(this.value * 1000) / 10 + '%')}"))
  ) %>%
  hc_title(text = "3 Probability Densities") %>%
  hc_add_series(
    name = "Normal", 
    type = "spline",
    #data = plot_data$normal,
    data = plot_data %>% select(x = value, y = normal) %>% list_parse(),
    marker = list(enabled = FALSE)
  ) %>%
  hc_add_series(
    name = "Exponential",
    type = "spline",
    #data = plot_data$exponential,
    data= plot_data %>% select(x = value, y = exponential) %>% list_parse(),
    marker = list(enabled = FALSE)
  ) %>%
  hc_add_series(
    name = "Uniform",
    type = "spline",
    #data = plot_data$uniform,
    data = plot_data %>% select(x = value, y = uniform) %>% list_parse(),
    marker = list(enabled = FALSE)
  )

```

<button id="test"> Click Me </button>

<script>

  $(document).ready(function() {

    var show_cdf = true;

    $("#test").click(function(){

      var chart=$("#densities_v2").highcharts();

      if (!show_cdf) {

        chart.yAxis[0].setTitle({text:"Cumulative Probability", redraw:false});
        uni_cdf = `r uni_cdf`;
        nor_cdf = `r nor_cdf`;
        exp_cdf = `r exp_cdf`;
        chart.series[0].setData(nor_cdf);
        chart.series[1].setData(exp_cdf);
        chart.series[2].setData(uni_cdf);
        $("#test").html("Show Density");

      } else {

        chart.yAxis[0].setTitle({text:"Probability Density", redraw:false});
        uni_pdf = `r uni_pdf`;
        nor_pdf = `r nor_pdf`;
        exp_pdf = `r exp_pdf`;
        chart.series[0].setData(nor_pdf);
        chart.series[1].setData(exp_pdf);
        chart.series[2].setData(uni_pdf);
        $("#test").html("Show CDF");

    };

    show_cdf = !show_cdf;

    }); 
  });

</script>

```{class.source="language-javascript"}

$(document).ready(function() {

  var show_cdf = false;

  $("#test").click(function(){

    var chart=$("#densities_v2").find(".highchart").highcharts();

    if (show_cdf) {

      chart.yAxis[0].setExtremes(0,null);
      chart.yAxis[0].setTitle({text:"Probability Density", redraw:false});
      uni_cdf = `r uni_cdf`;
      nor_cdf = `r nor_cdf`;
      exp_cdf = `r exp_cdf`;
      chart.series[0].setData(uni_cdf);
      chart.series[1].setData(nor_cdf);
      chart.series[2].setData(exp_cdf);

    } else {

      chart.yAxis[0].setExtremes(0,1);
      chart.yAxis[0].setTitle({text:"Cumulative Probability", redraw:false});

      $("#update_age_series").html("Show Density");

  };

  show_cdf = !show_cdf

  }); 
});

```