### 2019-08-16 04_highcharter-handbook/highcharter-handbook.Rmd
2nd work comp

---
title: "Highcharter Handbook"
author: "Tom Bishop"
date: "30 July 2019"
output: 
    html_document:
        toc: true
        toc_float: true        
        theme: sandstone
---

```{r setup, include=FALSE}

library(highcharter)
library(tidyverse)
library(forecast)
library(ggplot2)
knitr::opts_chunk$set(echo = TRUE)
```

## Intro

`highcharter` is an R implementation of the highcharts javascript graphing API through R's `htmlwidgets`. Most of the highcharts functionality is implemented through `highcharter` however the documentation is somewhat limited. This guide will provide examples on how to create and customise various graphs whilst providing some tips on how to think about the package that will help you build and debug your more ambitious charts. I'll try to show and explain (to the best of my knowledge) the most common graphs and customisation options I use in my day-to-day work.

Your most important resources should be the `highcharter` official website, and the highcharts api documentation which you should bookmark in your browser:

* [Highcharter Official Website](http://jkunst.com/highcharter/)
* [Highcharts API Documentation](https://api.highcharts.com/highcharts/)

As you become more a more proficient `highcharter` user and your charting requirements become more complicated you will need to understand more about the complexities of the highcharts API and you might need to learn + read a little javascript. This learning curve is actually the best thing about highcharter: basic charts are as easy to create as `ggplot`, whilst complex / beautiful chars are still achievable with enough thought and tinkering.

In the following examples I'll frequently be using the following two datasets:

* dplyr::starwars | the physical traits and info for characters in the starwars films
* dplyr::storms | hurricane timeseries data

## Basics

### Tidy Data

Most (~90%) of your day-to-day charts can be probably written with 2-3 lines of code using the `hchart` and `hcaes` functions. These functions allow you packages `ggplot` like syntax. That is if your

In `highcharter`, `ggplot`'s grammar of graphics aesthetic mapping is supported through the `hchart()` function.

A simple `ggplot` call might look like the following:

```{r message=FALSE, warning=FALSE}
ggplot(starwars, aes(x = name, y = height)) +
  geom_col()
```

This chart (let's be honest here) looks awful. For full disclosure `ggplot` has a number of functional advantages over highcharter as it has simple implementations of reasonably complex chart types. But highcharter creates quick, readable, nice looking default charts that can be presented in a usable publishing format (a html page) a thousand times better than ggplot.

Here's the highcharter version of this plot:

```{r message=FALSE, warning=FALSE}
hchart(starwars, "column", hcaes(x = name, y = height))
```

There's a number of things you might notice about this chart and how it differs from it's ggplot counterpart.

* 
* t thing is that if you can interact with it; hover over it to see a (relatively ugly) tooltip. Second is that it's dynamic, which you'll notice initially because there are more bars (star wars characters) than there are labels on the y-axis. To see an exaggerated version of this reduce the width of your browser window to see the chart react and replot as it tries to estimate the best way to present the data without it looking busy and cluttered; the number of characters shown on the y-axis will keep decreasing.

This is normally a great and useful feature of highcharts but it's something to keep track of as you plot.

### Default charts

## hchart()

The `hchart()` function is kind of an amazing function that has been written to create immediate logical graphs from various types of R objects. `hchart()` will detect the R object type and decide the best way to plot it. Here are some examples. Notice that the function call is exactly the same `hchart([object])` but the plot that's produced is very different.

Various other object types are supported which you can find [here](http://jkunst.com/highcharter/hchart.html)

### Numeric Vector :: Histogram

```{r}
starwars %>%
  pull(mass) %>%
  hchart(name = class(.))
```

### Density Object :: Density Plot

```{r}
starwars %>%
  filter(!is.na(mass)) %>%
  pull(mass) %>%
  density() %>%
  hchart(name = class(.))
```

### Forecast Object :: Timeseries / Forecast Chart

```{r}
storms %>%
  filter(name == "Emily") %>%
  pull(wind) %>%
  forecast() %>%
  hchart(name = class(.))
```